Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-06-21 doc/thesis/methods_results.tex
Almooooooost!

Two additional simulations were performed to address \del{potential sources of error}
\add{the possible impact of two types of model misspecification}. To consider
the effect of heterogeneity among nodes, we generated a network where half the
nodes were attached with power $\gls{alpha} = 0.5$ and the other half with power
$\gls{alpha} = 1.5$. The other parameters for this network were $\gls{N} = 5000$,
$\gls{I} = 1000$, and $\gls{m} = 2$. To investigate the effects of potential
sampling bias~\autocite{karcher2016quantifying}, we simulated a transmission
tree where the tips were sampled in a peer-driven fashion, rather than at
random. That is, the probability to sample a node was twice as high if any of
that node's network peers had already been sampled. The parameters of this
network were $\gls{N} = 5000$, $\gls{I} = 2000$, $\gls{m} = 2$, and
$\gls{alpha} = 0.5$.

\end{document}

@@

Trees simulated under different values of \gls{alpha} were visibly quite
distinct (\cref{fig:alphatrees}). In particular, higher values of \gls{alpha}
produce networks with a small number of highly connected nodes, which, once
infected, are likely to transmit to many other nodes. This results in a more
unbalanced, ladder-like structure in the phylogeny, compared to networks with
lower \gls{alpha} values. None of the other three parameters produced trees
that were as easily distinguished from each other
(\cref{fig:Itrees,fig:mtrees,fig:Ntrees,fig:Itrees}).  Sackin's index, which
measures tree imbalance, was significantly correlated with all four parameters
    (for $\alpha$, $I$, $m$, and $N$ respectively: Spearman's rho =
     0.85,
     \ensuremath{-0.12},
     \ensuremath{-0.13},
     0.09;
     $p$-values
     ${<}10^{-5}$,
     $0.003$,
     ${<}10^{-5}$,
     ${<}10^{-5}$).
The ratio of internal to terminal branch lengths was negatively correlated with
\gls{alpha} and \gls{I}, and positively correlated with \gls{m} and \gls{N}
  (Spearman's rho
    \ensuremath{-0.84},
    \ensuremath{-0.69},
    0.1,
    0.18;
  all $p < 10^{-5}$).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{kernel-alpha-tree.pdf}
  \caption[Simulated transmission trees under three different values of BA parameter $\alpha$]{
    Simulated transmission trees under three different values of BA parameter
    $\alpha$. Epidemics were simulated on \gls{BA} networks of 5000 nodes, with
    \gls{alpha} equal to 0.5, 1.0, or 1.5, until 1000 individuals were
    infected. Transmission trees were created by sampling 500 infected nodes.
    Higher \gls{alpha} values produced networks with a small number of
    highly-connected nodes, resulting in highly unbalanced, ladder-like trees.
  }
  \label{fig:alphatrees}
\end{figure}

\Cref{fig:kpca} shows \gls{kPCA} projections of the simulated trees onto the
first two principal components of the kernel matrix. The figure shows only the
simulations with 500-tip trees and 1000 infected nodes. The three \gls{alpha}
and \gls{I} values considered are well separated from each other in feature
space. On the other hand, the three \gls{N} values overlap significantly, and
the three \gls{m} values are virtually indistinguishable. Similar observations
can be made for other values of \gls{I} and the number of tips
(\cref{fig:alphakpca,fig:Nkpca,fig:Ikpca,fig:mkpca}). The values of \gls{I} and
\gls{N} separated more clearly with larger numbers of tips, and in the case of
\gls{N}, with larger epidemic sizes {\color{blue}(\cref{fig:Ikpca,fig:Nkpca})}.

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-kpca.pdf}
  \caption[Kernel-PCA projections of simulated trees under varying BA
           parameter values.]{
    Each parameter of the \gls{BA} model was individually varied to produce 300
    simulated trees. Kernel matrices were formed from all pairwise kernel
    scores among each set of 300 trees. The trees were projected onto the first
    two principal components of the kernel matrix calculated using \gls{kPCA}.
    All trees had 500 tips. The parameters not being varied were set to
    \gls{alpha} = 1, \gls{I} = 1000, \gls{m} = 2, and \gls{N} = 5000. The tree
    kernel meta-parameters were $\lambda = 0.3$ and $\sigma = 4$.
  }
  \label{fig:kpca}
\end{figure}

Accuracy of the \gls{kSVR} classifiers varied based on the parameter being
tested (\cref{fig:rsquared}, left). Classifiers based on two other tree
statistics, the \gls{nltt} and Sackin's index, generally exhibited worse
performance than the tree kernel, although the magnitude of the disparity
varied between the parameters (\cref{fig:rsquared}, centre and right). The
results were largely robust to variations in the tree kernel meta-parameters
$\lambda$ and $\sigma$, although accuracy varied between different epidemic and
sampling scenarios
(\cref{fig:alphacrossv,fig:mcrossv,fig:Icrossv,fig:Ncrossv}).

When classifying $\alpha$, the \gls{kSVR} classifier had an average $R^2$ of 
    0.92,
compared to 
    0.56
for the \gls{nltt}-based SVR, and
    0.75
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify the $m$ parameter in any epidemic scenario, with average
$R^2$ values of 
  0.12 for \gls{kSVR},
  0.01 for the \gls{nltt}, and
  0.06
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVR} was slightly higher
on 1000-tip trees 
    (average $R^2$ 
     0.01,
     0.11,
     0.32
     for 100, 500, and 1000 tips respectively).

The accuracy of classifiers for $I$ varied significantly with the number of
tips in the tree. For 100-tip trees, the average $R^2$ values were
  0.7,
  0.55, and
  0.02
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  0.93,
  0.83, and
  0.07.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  0.08
for the smallest epidemic and smallest sample size, to
  0.82
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  0.01
to
  0.54.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  0.03
and little variation between scenarios.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
  \caption[Cross-validation accuracy of kernel-SVR, nLTT-based SVR, and
  Sackin's index regression classifiers for BA model parameters.]{
      Cross-validation accuracy of kernel-SVR classifier (left), SVR classifier
      using \gls{nltt} (centre), and linear regression using Sackin's index
      (right) for \gls{BA} model parameters. Kernel meta-parameters were set to
      $\lambda = 0.3$ and $\sigma = 4$. Each point was calculated based on 300
      simulated transmission trees over networks with three different values of
      the parameter being tested, assuming perfect knowledge of the other
      parameters. Vertical lines are empirical 95\% confidence intervals based
      on 1000 two-fold cross-validations. {\color{blue}\uline{ The classifiers
      for \gls{I} were not evaluated with 1000-tip trees, because one of the
      tested \gls{I} values was 500, and it is not possible to sample a tree of
      size 1000 from 500 infected individuals.}}
  }
  \label{fig:rsquared}
\end{figure}

@@

{\color{red}\sout{The accuracy of grid search estimates largely paralleled that
of the \gls{kSVR} classifiers.}}\Cref{fig:gridest} shows point estimates and
{\color{blue}\uline{50\%}} and 95\% highest density intervals for each of the
\gls{BA} parameters, for one replicate experiment with 500-tip trees. Plots
showing the point estimates for all replicates can be found in
\cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all parameters
except $m$, the error of point estimates was negatively correlated with the
number of sampled tips in the tree (for \gls{alpha}, \gls{I}, and \gls{N}
respectively: Spearman's $\rho$ = \ensuremath{-0.22},
\ensuremath{-0.51}, \ensuremath{-0.16}; $p$-values
    $4\!\times\!10^{-4}$,
    ${<}10^{-5}$,
    $0.01$).
The 95\% highest density intervals obtained for all parameters were extremely
wide, occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

{\color{blue}\uline{Across all replicates, $R^2$ values for the correlations
between the estimated and true values were
    $0.91$,
    $0.91$,
    $0.25$, and
    $0.54$
for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively. The mean absolute
errors of the point estimates were
    $0.14$,
    $310$,
    $1.31$, and
    $2419$,
representing
    $7$\%,
    $8$\%,
    $26$\%, and
    $17$\%
of the respective grids.}

\uline{Qualitatively, each of the parameters exhibited weak identifiability
within particular sections of the grid} (\cref{fig:gridest}). \uline{The 50\%
\glspl{HDI} for \gls{alpha} were nearly identical for the values $\gls{alpha}
\leq 0.5$ (on average [X-X]) and for $\gls{alpha} \geq 1.5$ ([X-X]).
For \gls{I}, similar 50\% \gls{HPD} were observed for $\gls{I} \leq 1500$
(average [X-X]) and for $\gls{I} > 3000$ ([X-X]). For \gls{m}, both the 50\%
and 95\% \glspl{HDI} were generally identical for $\gls{m} > 2$ (average [X-X]
for 50\% and [X-X] for 95\% \gls{HDI}). We did not observe a similar pattern
for \gls{N}, however the boundaries of the \gls{HDI} did not always
monotonically increase with the actual value.}

\Cref{fig:gridalpha,fig:gridI,fig:gridm,fig:gridN} \uline{show kernel score
distributions of kernel scores along the grid for each parameter. The
distributions for some values, such as $\gls{alpha} = 1.25$, $\gls{I} = 500$
and $4500$, $\gls{m} = 1$, and $\gls{N} = 1000$, exhibited distinct peaks
around the true value. This indicates that these values produce distinctively
shaped trees that can be identified with the tree kernel, when the other
parameter values are fixed and known. However, for the majority of values of
each parameter, the score distributions were fairly flat around the true value. 
This means there is a range of values which produce similarly shaped trees, and
the parameter is less identifiable within that range. The exception was
\gls{I}, whose score distributions exhibited a more or less rounded shape with
the highest point near the true value. }}

{\color{red}\sout{The \gls{alpha} parameter was the most accurately estimated,
with point estimates having an average deviation of 
    0.14
from the true value, on a grid from 0 to 2. The error of point estimates varied
significantly between true values of \gls{alpha}
    (one-way \gls{ANOVA}, $p$ ${<}10^{-5}$). In
particular, errors were lower for the values \gls{alpha} = 1.0 and 1.25 than
for the other values
    (average errors 
    0.03
    for \gls{alpha} = 1.0 or 1.5 vs.
    0.17
    for \gls{alpha} $\neq$ 1.0 or 1.5),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
}\cref{fig:gridptalpha}).
\sout{These two values exhibited different qualitative behaviour than the other
values in terms of the distribution of kernel scores along the grid
}(\cref{fig:gridalpha}). \sout{In particular, there was a pronounced peak in scores
around the true value, in contrast to the other values where the scores were
flat around the true value. The effect was most obvious for the value
\gls{alpha} = 1.25.}}

%The average absolute error of the point estimates for \gls{I} was 
%    d["I", round(mean(error))] individuals,
%on a grid of 500 to 5000, and these errors differed between true values of
%\gls{I}
%    (one-way \gls{ANOVA}, $p pp(av["I", p], eq=TRUE)$).
%The errors for $2000 \leq I \leq 3000$ were higher than those for the other
%values
%    (average errors
%     d["I", .SD[true_value >= 2000 & true_value <= 3000, round(mean(error))]]
%     for $2000 \leq I \leq 3000$ vs.
%     d["I", .SD[true_value < 2000 | true_value > 3000, round(mean(error))]]
%     for $I < 2000$ or $I > 3000$),
%and this difference was significant
%    (Wilcoxon rank-sum test, $p pp(I.test$p.value, eq=TRUE)$,
%     \cref{fig:gridptI}).
%Kernel score distributions for all test values exhibited a similar rounded
%shape (\cref{fig:gridI}). 
%
%The average error for \gls{m} was
%    d["m", round(mean(error), 2)] edges per vertex,
%on a grid from 1 to 6. The error varied significantly between the true values
%of \gls{m} 
%    (one-way \gls{ANOVA}, $p pp(av["m", p], eq=TRUE)$).
%Errors for the value \gls{m} = 1 were lower than the other values
%    (average errors
%    d["m", .SD[true_value == 1, round(mean(error), 2)]]
%    for $m = 1$ vs.
%    d["m", .SD[true_value > 1, round(mean(error), 2)]]
%    for $m > 1$),
%and this difference was significant
%    (Wilcoxon rank-sum test, $p pp(m.test$p.value, eq=TRUE)$,
%     \cref{fig:gridptm}).
%The value $m = 1$ causes the network to take on a distinct shape relative to
%higher \gls{m} values, namely a tree (\ie there are no cycles, see
%\cref{subsec:treeshape}). The kernel score distribution had a peak at $m = 1$
%when this was the true value, and a valley at $m = 1$ when the true value of
%$m$ was greater that 1 (\cref{fig:gridm}).
%
%The average error for \gls{N} was 
%    d[parameter == "N", round(mean(error))] individuals,
%on a grid from 1000 to 15000, and was varied significantly with the true value
%of \gls{N}
%    (one-way \gls{ANOVA}, $p pp(av["N", p], eq=TRUE)$).
%The errors were lower for $N \leq 3000$
%    (average errors
%     d["N", .SD[true_value <= 3000, round(mean(error))]]
%     for $N \leq 3000$ vs.
%     d["N", .SD[true_value > 3000, round(mean(error))]]
%     for $N > 3000$),
%and this difference was significant
%    (Wilcoxon rank-sum test, $p pp(N.test$p.value, eq=TRUE)$,
%     \cref{fig:gridptN}).
%The kernel score distribution had a peak at $N = 1000$ when this was the true
%value, and a valley there otherwise (\cref{fig:gridN}). Except for this valley,
%the distributions were flat for $N > 3000$.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{gridsearch-example}
  \caption[
    Grid search estimates of \gls{BA} model parameters for one replicate
    experiment with trees of size 500.
  ]{
    Grid search estimates of \gls{BA} model parameters for one replicate
    experiment with trees of size 500. Point estimates (dots), 95\% highest
    density intervals (lines), and 50\% highest density intervals (notches) for
    each \gls{BA} model parameter, obtained using grid search. Networks and
    transmission trees were simulated over a grid of values for each parameter
    while holding the others fixed. For a subset of the grid values ($x$-axis),
    test networks and trees were created and compared to each tree on the grid
    using the tree kernel. The kernel scores along the grid were normalized to
    resemble a probability distribution, from which the mode and highest
    density interval were calculated. 
  } 
  \label{fig:gridest}
\end{figure}

@@

\Cref{fig:abcptm2} shows {\color{red}\sout{\gls{MAP}}}
{\color{blue}\uline{posterior mean}} point estimates of the BA model parameters
obtained with ABC on simulated data. The estimates shown
correspond only to the simulations where the $m$ parameter was set to 2,
however the results for $m = 3$ and $m = 4$ were similar
(\cref{fig:abcptm3,fig:abcptm4}). Average boundaries of 95\% \gls{HPD}
intervals are given in \cref{tab:abchpd}.

{\color{blue}
\uline{Across all simulations, the median [\gls{IQR}] absolute errors of the parameter 
estimates obtained with \software{netabc} were
    0.11 
    [0.03 - 
    0.25]
for \gls{alpha},
    492 
    [294 - 
    782]
for \gls{I},
    1 
    [0 - 
    1]
for \gls{m}, and
    4153 
    [3660 - 
    4489]
for \gls{N}. These errors comprised, respectively,
    6\%,
    11\%,
    17\%, and
    29\%
of the regions of nonzero prior density. Average \gls{HPD} interval widths were
    0.68,
    2454,
    3.01, and
    12046,
representing 
    34\%,
    55\%,
    50\%, and
    83\%
of the nonzero prior density regions. Point estimates of \gls{I} were upwardly
biased: \gls{I} was overestimated in
    69
out of 
    72
simulations
    (96\%).
The estimates for \gls{m} and \gls{N} were similar across all simulations
(median [IQR] point estimates
    3
    [3 - 
    3]
and 
    9153
    [8660 - 
    9489])
regardless of the true values of any of the \gls{BA} parameters.}

\uline{To analyse the effects of the true parameter values on the
accuracy our estimates of \gls{alpha} and \gls{I}, we fitted one \glspl{GLM}
for each of these two parameters, with error rate as the dependent variable and
the true parameter values as independent variables. Since the estimates of
\gls{m} and \gls{N} were roughly equal across all simulations}
(\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}), \uline{\glspl{GLM} were not
fitted for these parameters. The estimated coefficients are shown in
}\cref{tab:glmalpha,tab:glmI}. \uline{For the parameters \gls{alpha}, \gls{I},
and \gls{N}, the \gls{GLM} were fitted using the inverse link function. That
is, if $p$ is the true value of the parameter and $\hat{p}$ is a random
variable representing our estimate of the parameter, the \gls{GLM} posits a
relationship of the form }
\[
    \E(|p-\hat{p}|) = (\beta_0 + \beta_\alpha \alpha + \beta_I I + \beta_m m)^{-1},
\]
\uline{where the $\beta$'s are coefficients to be fitted. If the true value
\gls{alpha}, say, is increased by one, the \emph{inverse} of the expected
absolute error will increase by $\beta_\alpha$. If $\beta_\alpha$ is positive,
it means that the absolute error decreases as the true value of \gls{alpha}
increases.}}

>>>>

Two further simulations were performed to address \del{potential sources of
error} \add{the possible impact of two types of model misspecification}. To
consider the effect of heterogeneity among nodes, we generated a network where
half the nodes were attached with power $\gls{alpha} = 0.5$ and the other half
with power $\gls{alpha} = 1.5$. The other parameters for this network were
$\gls{N} = 5000$, $\gls{I} = 1000$, and $\gls{m} = 2$. To investigate the
effects of potential sampling bias~\autocite{karcher2016quantifying}, we
simulated a transmission tree where the tips were sampled in a peer-driven
fashion, rather than at random. That is, the probability to sample a node was
twice as high if any of that node's network peers had already been sampled. The
parameters of this network were $\gls{N} = 5000$, $\gls{I} = 2000$, $\gls{m} =
2$, and $\gls{alpha} = 0.5$.

\add{To assess the impact of the \gls{SMC} settings on \software{netabc}'s
accuracy, we ran \software{netabc} twice on the same simulated transmission
tree. For the first run, the \gls{SMC} settings were the same as in the other
simulations: 1000 particles, 5 simulated transmission trees per particle, and
\gls{alphaess} = 0.95. The second run was performed with 2000 particles, 10
simulated transmission trees per particle, and \gls{alphaess} = 0.99. To
investigate the extent to which errors in the estimated \gls{BA} parameters
were due to true features of the posterior, rather than an inaccurate \gls{ABC}
approximation, we performed marginal estimation for one set of parameter
values. Each combination of 1, 2, or 3 model parameters (14 combinations total)
was fixed to their known values, and the remaining parameters were estimated
with \software{netabc}. The parameter values were $\gls{alpha} = 0.0$, $\gls{m}
= 2$, $\gls{I} = 2000$, and $\gls{N} = 5000$.}

@@

Trees simulated under different values of \gls{alpha} were visibly quite
distinct (\cref{fig:alphatrees}). In particular, higher values of \gls{alpha}
produce networks with a small number of highly connected nodes, which, once
infected, are likely to transmit to many other nodes. This results in a more
unbalanced, ladder-like structure in the phylogeny, compared to networks with
lower \gls{alpha} values. None of the other three parameters produced trees
that were as easily distinguished from each other
(\cref{fig:Itrees,fig:mtrees,fig:Ntrees,fig:Itrees}).  Sackin's index, which
measures tree imbalance, was significantly correlated with all four parameters
    (for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively: Spearman's rho =
    0.85,
     \ensuremath{-0.12},
     \ensuremath{-0.13},
     0.09;
     $p$-values
     ${<}10^{-5}$,
     $0.003$,
     ${<}10^{-5}$,
     ${<}10^{-5}$).
The ratio of internal to terminal branch lengths was negatively correlated with
\gls{alpha} and \gls{I}, and positively correlated with \gls{m} and \gls{N}
  (Spearman's rho
    \ensuremath{-0.8},
    \ensuremath{-0.69},
    0.09,
    0.17;
all $p < 10^{-5}$).

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{kernel-alpha-tree.pdf}
    \caption[
        Simulated transmission trees under three different values of
        preferential attachment power (\gls{alpha}) parameter of \acrshort{BA}
        model.
    ]{
        Simulated transmission trees under three different values of
        preferential attachment power (\gls{alpha}) parameter of \acrshort{BA}
        model. Epidemics were simulated on \gls{BA} networks of 5000 nodes,
        with \gls{alpha} equal to 0.5, 1.0, or 1.5, until 1000 individuals were
        infected. Transmission trees were created by randomly sampling 500
        infected nodes. Higher \gls{alpha} values produced networks with a
        small number of highly-connected nodes, resulting in highly unbalanced,
        ladder-like trees.
    }
  \label{fig:alphatrees}
\end{figure}

\Cref{fig:kpca} shows \gls{kPCA} projections of the simulated trees onto the
first two principal components of the kernel matrix. The figure shows only the
simulations with 500-tip trees and 1000 infected nodes. The three \gls{alpha}
and \gls{I} values considered are well separated from each other in the feature
space \add{mapped to by the tree kernel}. On the other hand, the three \gls{N}
values overlap significantly, and the three \gls{m} values are virtually
indistinguishable. Similar observations can be made for other values of \gls{I}
and the number of tips (\cref{fig:alphakpca,fig:Nkpca,fig:Ikpca,fig:mkpca}).
The values of \gls{I} and \gls{N} separated more clearly with larger numbers of
tips, and in the case of \gls{N}, with larger epidemic sizes
\add{(\cref{fig:Ikpca,fig:Nkpca})}.

\begin{figure}[ht]
    \centering
    \includegraphics{kernel-kpca.pdf}
    \caption[
        Kernel-\gls{PCA} projections of simulated trees under varying \gls{BA}
        parameter values.
    ]{
        Each parameter of the \gls{BA} model was individually varied to produce
        300 simulated trees with 500 tips each. Kernel matrices were formed
        from all pairwise kernel scores among each set of 300 trees. The trees
        were projected onto the first two principal components of the kernel
        matrix calculated using \gls{kPCA}. The other parameters, which were
        not varied, were set to $\gls{alpha} = 1$, $\gls{I} = 1000$, $\gls{m} =
        2$, and $\gls{N} = 5000$. The tree kernel meta-parameters were
        $\gls{lambda} = 0.3$ and $\gls{sigma} = 4$.
  }
  \label{fig:kpca}
\end{figure}

The accuracy of each classifier \add{on all \gls{BA} parameters is shown in}
\del{varied based on the parameter being tested} \cref{fig:rsquared}.
Classifiers based on \del{two other tree statistics,} the \gls{nltt} and
Sackin's index generally exhibited worse performance than the tree kernel,
although the magnitude of the disparity varied between the parameters
(\cref{fig:rsquared}, centre and right). The results were largely robust to
variations in the tree kernel meta-parameters \gls{lambda} and \gls{sigma},
although accuracy varied between different epidemic and sampling scenarios
(\cref{fig:alphacrossv,fig:mcrossv,fig:Icrossv,fig:Ncrossv}). \add{For all
parameters except \gls{m}, the absolute number of tips in the tree had a much
greater impact on accuracy than the proportion of infected individuals these
tips represented. However, \gls{m}, both the number and proportion of sampled
tips had a strong impact or the accuracy of the \gls{kSVR}
(\cref{fig:mcrossv}).}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
    \caption[
        Cross-validation accuracy of kernel-SVR, nLTT-based SVR, and Sackin's
        index regression classifiers for BA model parameters.
    ]{
        Cross-validation accuracy of kernel-SVR classifier (left), SVR
        classifier using \gls{nltt} (centre), and linear regression using
        Sackin's index (right) for \gls{BA} model parameters. Kernel
        meta-parameters were set to $\gls{lambda} = 0.3$ and $\gls{sigma} = 4$.
        Each point was calculated based on 300 simulated transmission trees
        over networks with three different values of the parameter being
        tested, assuming perfect knowledge of the other parameters. Vertical
        lines are empirical 95\% confidence intervals based on 1000 two-fold
        cross-validations. \add{The classifiers for \gls{I} were not evaluated
        with 1000-tip trees, because one of the tested \gls{I} values was 500,
        and it is not possible to sample a tree of size 1000 from 500 infected
        individuals.}
    }
    \label{fig:rsquared}
\end{figure}

The \gls{kSVR} classifier for \gls{alpha} had an average $R^2$ of 
    0.92,
compared to 
    0.56
for the \gls{nltt}-based SVR, and
    0.75
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify \gls{m} in any epidemic scenario, with average $R^2$ values
of 
  0.12 for \gls{kSVR},
  0.01 for the \gls{nltt}, and
  0.06
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVR} was slightly higher
on 1000-tip trees 
    (average $R^2$ 
     0.01,
     0.11,
     0.32
     for 100, 500, and 1000 tips respectively).

The accuracy of classifiers for $I$ \del{varied significantly
with}\add{appeared to be strongly influenced by} the number of tips in the
tree. For 100-tip trees, the average $R^2$ values were
  0.7,
  0.55, and
  0.02
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  0.93,
  0.83, and
  0.07.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  0.08
for the smallest epidemic and smallest sample size, to
  0.82
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  0.01
to
  0.54.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  0.03
and little variation between scenarios.

@@

\del{The accuracy of grid search estimates largely paralleled that of the
\gls{kSVR} classifiers.} \Cref{fig:gridest} shows point estimates and
\add{50\%} and 95\% highest density intervals for each of the \gls{BA}
parameters, for one replicate experiment with 500-tip trees. Plots showing the
point estimates for all replicates can be found in
\cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all parameters
except \gls{m}, the error of point estimates was negatively correlated with the
number of sampled tips in the tree (for \gls{alpha}, \gls{I}, and \gls{N}
respectively: Spearman's $\rho$ = 
    \ensuremath{-0.22},
    \ensuremath{-0.51}, 
    \ensuremath{-0.16}; 
$p$-values
    $4\!\times\!10^{-4}$,
    ${<}10^{-5}$,
    $0.01$).
The 95\% highest density intervals obtained for all parameters were extremely
wide, occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{gridsearch-example}
    \caption[
        Grid search estimates of \gls{BA} model parameters for one replicate
        experiment with trees of size 500.
    ]{
        Grid search estimates of \gls{BA} model parameters for one replicate
        experiment with trees of size 500. Point estimates (dots), 95\%
        \glspl{HDI} (lines), and 50\% \glspl{HDI} (notches) for each \gls{BA}
        model parameter, obtained using grid search. Networks and transmission
        trees were simulated over a grid of values for each parameter while
        holding the others fixed \add{to known values}. For a subset of the
        grid values ($x$-axis), test networks and trees were created and
        compared to each tree on the grid using the tree kernel. The kernel
        scores along the grid were normalized to resemble a probability
        distribution, from which the mode and highest density interval were
        calculated.
    } 
    \label{fig:gridest}
\end{figure}

\add{Across all replicates, $R^2$ values for the correlations between the
estimated and true values were
    $0.91$,
    $0.91$,
    $0.25$, and
    $0.54$
for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively. The mean absolute
errors of the point estimates were
    $0.14$,
    $310$,
    $1.31$, and
    $2419$,
representing
    $7$\%,
    $8$\%,
    $26$\%, and
    $17$\%
of the respective grids. For \gls{I} and \gls{N}, relative errors may be more
appropriate to consider. An overestimate of 100 individuals would be very
misleading if the true population was only of size 100, but almost negligable
in a population of size 5000. The mean relative errors for \gls{I} and \gls{N}
respectively were
    $18$\% and 
    $31$\%.}

\add{Qualitatively, \gls{alpha} and \gls{I} exhibited weak identifiability
within particular sections of the grid (\cref{fig:gridest}). The 50\%
\glspl{HDI} for \gls{alpha} were similar for the values $\gls{alpha}
\leq 0.5$ (on average 
    0.02 -
    0.84)
and for $\gls{alpha} \geq 1.5$ 
    (1.26 -
    2).
For \gls{I}, similar 50\% \gls{HPD} were observed for $\gls{I} \leq 1500$
(average 
    [650 -
     2846])
and for $\gls{I} > 3000$ 
    ([2596 -
      4802]).
No similar patterns were observed for \gls{m} or \gls{N} (although the 50\%
\glspl{HDI} appear to be identical for $\gls{m} > 2$ in \cref{fig:gridexpt},
this was not consistent across replicates).}

\add{\Cref{fig:gridalpha,fig:gridI,fig:gridm,fig:gridN} show kernel score
distributions of kernel scores along the grid for each parameter. The
distributions for some values, such as $\gls{alpha} = 1.25$, $\gls{I} = 500$
and $4500$, $\gls{m} = 1$, and $\gls{N} = 1000$, exhibited distinct peaks
around the true value. This indicates that these values produce distinctively
shaped trees that can be identified with the tree kernel, when the other
parameter values are fixed and known. However, for the majority of values of
each parameter, the score distributions were fairly flat around the true value. 
This means there is a range of values which produce similarly shaped trees, and
the parameter is less identifiable within that range. The exception was
\gls{I}, whose score distributions exhibited a more or less rounded shape with
the highest point near the true value.}

\del{The \gls{alpha} parameter was the most accurately estimated,
with point estimates having an average deviation of 
    0.14
from the true value, on a grid from 0 to 2. The error of point estimates varied
significantly between true values of \gls{alpha}
    (one-way \gls{ANOVA}, $p$ ${<}10^{-5}$). In
particular, errors were lower for the values \gls{alpha} = 1.0 and 1.25 than
for the other values
    (average errors 
    0.03
    for \gls{alpha} = 1.0 or 1.5 vs.
    0.17
    for \gls{alpha} $\neq$ 1.0 or 1.5),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
\cref{fig:gridptalpha}). These two values exhibited different qualitative
behaviour than the other values in terms of the distribution of kernel scores
along the grid (\cref{fig:gridalpha}). In particular, there was a pronounced
peak in scores around the true value, in contrast to the other values where the
scores were flat around the true value. The effect was most obvious for the
value \gls{alpha} = 1.25.}

@@

\Cref{fig:abcptm2} shows \del{MAP} \add{posterior mean} point estimates of the
\gls{BA} model parameters obtained with \gls{ABC} on simulated data. The
estimates shown correspond only to the simulations where \gls{m} was set to 2,
however the results for $\gls{m} = 3$ and $\gls{m} = 4$ were similar
(\cref{fig:abcptm3,fig:abcptm4}). Average boundaries of 95\% \gls{HPD}
intervals are given in \cref{tab:abchpd}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{abc-point-estimate-m2}
    \vspace{6pt}
    \caption[
        Posterior mean point estimates for \gls{BA} model parameters obtained
        by running \software{netabc} on simulated data, for simulations with
        $\gls{m} = 2$.
    ]{
        Posterior mean point estimates for \gls{BA} model parameters obtained
        by running \software{netabc} on simulated data, for simulations with
        $\gls{m} = 2$. Dashed lines indicate true values. (A) Estimates of
        \gls{alpha} and \gls{I} which were varied in these simulations against
        known values. (B) Estimates of \gls{m} and \gls{N} which were held
        fixed in these simulations at the values $\gls{m} = 2$ and 
        $\gls{N} = 5000$.
    }
    \label{fig:abcptm2}
\end{figure}

\begin{table*}[ht]
    \centering
    \input{\tablepath/abc-hpd.tex}
    \caption[
        Average posterior mean point estimates and 95\% \glspl{HDI} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data.
    ]{
        Average posterior mean point estimates and 95\% \glspl{HDI} interval
        widths for \gls{BA} model parameter estimates obtained with
        \software{netabc} on simulated data. Three transmission trees were
        simulated under each combination of the listed parameter values, and
        the parameters were estimated with \gls{ABC} without training.
    }
    \label{tab:abchpd}
\end{table*}

\add{Across all simulations, the median [\gls{IQR}] absolute errors of the parameter 
estimates obtained with \software{netabc} were
    0.11 
    [0.03 - 
    0.25]
for \gls{alpha},
    492 
    [294 - 
    782]
for \gls{I},
    1 
    [0 - 
    1]
for \gls{m}, and
    4153 
    [3660 - 
    4489]
for \gls{N}. These errors comprised, respectively,
    6\%,
    11\%,
    17\%, and
    29\%
of the regions of nonzero prior density. For \gls{I} and \gls{N}, relative
errors were
    38\%
    [20 - 
    50\%]
and
    83\%
    [73 - 
    90\%].
Average \gls{HPD} interval widths were
    0.68,
    2454,
    3.01, and
    12046,
representing 
    34\%,
    55\%,
    50\%, and
    83\%
of the nonzero prior density regions. Point estimates of \gls{I} were upwardly
biased: \gls{I} was overestimated in
    69
out of 
    72
simulations
    (96\%).
The estimates for \gls{m} and \gls{N} were similar across all simulations
(median [IQR] point estimates
    3
    [3 - 
    3]
and 
    9153
    [8660 - 
    9489])
regardless of the true values of any of the \gls{BA} parameters.}
